---
title: "Chapter_2_analysis"
author: "Margarita Orlova"
date: "February 8, 2020"
output:
  pdf_document: default
  html_document: default
---

First we load MLE estimates along with genotypes for top  235 genes and regression model outputs.

```{r}
library(rlist)
estimates<-list.load("C:/Users/Moonkin/Documents/GitHub/Thesis_single_RNA/analysis/estimates_for_235_genes.rds")
regression<-read.table("C:/Users/Moonkin/Documents/GitHub/Thesis_single_RNA/analysis/regression_for_235_genes.txt", header = T, sep = "\t")
```

First we can take a look at quantile-quantile plots for $k_{on}$,$k_{off}$,$k_{r}$ p-values 

```{r}
par(mfrow=c(1,3))
U=seq(0, 1, length.out = 236)
U=U[2:236]
#plot(U,sort(regression$k_on_slope_p))
#lines(U,U)
plot(-log(U,base=10),-log(sort(regression$k_on_slope_p),base=10), main="k_on")
lines(-log(U,base=10),-log(U,base=10))

plot(-log(U,base=10),-log(sort(regression$k_off_slope_p),base=10), main="k_off")
lines(-log(U,base=10),-log(U,base=10))

plot(-log(U,base=10),-log(sort(regression$k_r_slope_p),base=10), main="k_r")
lines(-log(U,base=10),-log(U,base=10))
```

We apply Benjamini-Hochberg procedure to correct for FDR.


```{r}
regression[p.adjust(regression$k_on_slope_p, method = "BH" , n = length(regression$k_off_slope_p))<0.05,]
regression[p.adjust(regression$k_off_slope_p, method = "BH" , n = length(regression$k_off_slope_p))<0.05,]
regression[p.adjust(regression$k_r_slope_p, method = "BH" , n = length(regression$k_off_slope_p))<0.05,]
```


```{r, echo=FALSE}
#BH_1=regression[order(regression$k_on_slope_p),]
#critical_p=(seq(1,235, by=1)/235)*0.05
#BH_1[BH_1$k_on_slope_p<critical_p]
```

Here the scatterplot of estimates vs genotype for ENSG00000197728 gene, the only one that with all slopes coefficients being significant at 0.05 level after BH adjustment.

```{r}
ENSG00000197728=estimates[[2]]

par(mfrow=c(1,3))
model1<- lm(k_on~genotype, data=ENSG00000197728)
model2<- lm(k_off~genotype, data=ENSG00000197728)
model3<- lm(k_r~genotype, data=ENSG00000197728)

plot(ENSG00000197728$genotype,ENSG00000197728$k_on, xlab="genotype", ylab="k_on estimate", main= "k_on scatterplot by genotype")
abline(model1, col="red")

plot(ENSG00000197728$genotype,ENSG00000197728$k_off, xlab="genotype", ylab="k_off estimate", main= "k_off scatterplot by genotype")
abline(model2, col="red")

plot(ENSG00000197728$genotype,ENSG00000197728$k_r, xlab="genotype", ylab="k_r estimate", main= "k_r scatterplot by genotype")
abline(model3, col="red")
```







