---
title: "MLE estimation with noise parameter"
author: "Margarita Orlova"
date: "January 30, 2020"
output: html_document
---

In this section we  have tried to implement $s_i$ noise parameter as in Kim and Marioni paper.

The model is $x|k_r,p \sim  Poisson (s_ik_rp_i)$ where $p|k_{on},k_{off}\sim Beta(k_{on},k_{off})$ and $s_i$ is a total number of molecules which were sequenced in a cell, which captures noise (higher $s_i$ means lower noise)


```{r, echo=FALSE}
library(reticulate)
library(robustbase)
source_python("C:/Users/Moonkin/Documents/GitHub/Thesis_single_RNA/analysis/MLE.py")
```


# One dataset example

First, we generate one dataset  and estimate MLE without noise variable:

```{r}
k_on=as.numeric(0.1)
k_off=as.numeric(0.1)
kr=as.numeric(100)
n_cells=as.integer(100)
x=generate_data(k_on,k_off,kr,n_cells)
```


```{r}
est=MaximumLikelihood(x)
est
```

Then, we implement MLE calculation with $s_i$ parameter included.

```{r}
source_python("C:/Users/Moonkin/Documents/GitHub/Thesis_single_RNA/analysis/MLE_with_si.py")
```

With $s_i$=1 we should receive the same estimators as without $s_i$

```{r}
si=matrix(1, nrow = 100, ncol = 1)
est=MaximumLikelihood_with_si(x,si)
est
```

With $s_i$=2, our $k_r$ estimate drops twice, while $k_{on}$ and $k_{off}$ are not really affected by this change.

```{r}
si=matrix(2, nrow = 100, ncol = 1)
est=MaximumLikelihood_with_si(x,si)
est
```

We also try using different $s_i$ values as this is what we will see in real data:

```{r}
si=matrix(1:5, nrow = 100, ncol = 1)
est=MaximumLikelihood_with_si(x,si)
est
```


# Multiple dataset results

Now we will also test model with $s_i$ vs the model without it:

Initial MLE calculations:

```{r,echo=FALSE}
#True k_on=k_off=0.1; k_r=100
n_estimates=1000
k_on=as.numeric(0.1)
k_off=as.numeric(0.1)
kr=as.numeric(100)
n_cells=as.integer(100)
est=matrix(, nrow = n_estimates, ncol = 3)
for (i in 1:n_estimates){
  x=generate_data(k_on,k_off,kr,n_cells)
  est[i,]=MaximumLikelihood(x)
}
hist(log(est[,1]),breaks=30, main="MLE estimate of k_on parameter,1000 runs, true k_on=0.1, log scale", xlab="k_on MLE estimate")
paste("true k_on is 0.1;","k_on mean",round(mean(est[,1],na.rm=TRUE),3),"k_on median",round(median(est[,1],na.rm=TRUE),3))
hist(log(est[,2]),breaks=30, main="MLE estimate of k_off parameter,1000 runs, true k_off=0.1, log scale", xlab="k_off MLE estimate")
paste("true k_off is 0.1;","k_off mean",round(mean(est[,2],na.rm=TRUE),3),"k_off median",round(median(est[,2],na.rm=TRUE),3))
hist(log(est[,3]),breaks=30, main="MLE estimate of k_r parameter,1000 runs,true k_r=100, log scale", xlab="k_r MLE estimate")
paste("true k_r is 100;","k_r mean",round(mean(est[,3],na.rm=TRUE),3),"k_r median",round(median(est[,3],na.rm=TRUE),3))
```


MLE calculations with $s_i$=2 for all cells:

```{r,echo=FALSE}
#True k_on=k_off=0.1
n_estimates=1000
k_on=as.numeric(0.1)
k_off=as.numeric(0.1)
kr=as.numeric(100)
n_cells=as.integer(100)
si=matrix(2, nrow = 100, ncol = 1)
est=matrix(, nrow = n_estimates, ncol = 3)
for (i in 1:n_estimates){
  x=generate_data(k_on,k_off,kr,n_cells)
  est[i,]=MaximumLikelihood_with_si(x,si) 
}
hist(log(est[,1]),breaks=30, main="MLE estimate of k_on parameter,1000 runs, true k_on=0.1, log scale", xlab="k_on MLE estimate")
paste("true k_on is 0.1;","k_on mean",round(mean(est[,1],na.rm=TRUE),3),"k_on median",round(median(est[,1],na.rm=TRUE),3))
hist(log(est[,2]),breaks=30, main="MLE estimate of k_off parameter,1000 runs, true k_off=0.1, log scale", xlab="k_off MLE estimate")
paste("true k_off is 0.1;","k_off mean",round(mean(est[,2],na.rm=TRUE),3),"k_off median",round(median(est[,2],na.rm=TRUE),3))
hist(log(est[,3]),breaks=30, main="MLE estimate of k_r parameter,1000 runs,true k_r=100, log scale", xlab="k_r MLE estimate")
paste("true k_r is 100;","k_r mean",round(mean(est[,3],na.rm=TRUE),3),"k_r median",round(median(est[,3],na.rm=TRUE),3))
```


# Real data

We also estimate MLE with size parameter on the real data.

We use top gene with smallest p_beta parameter and obtain $s_i$ for each cell from scqtl-annotation.txt.gz, mol_hs field

The MLE estimation without noise can be found here https://margareth2407.github.io/Thesis_single_RNA/Real_data_filter.html

```{r, echo=FALSE}
mol_hs <-read.table("D:/Uchicago/Thesis/Real_data/scqtl-annotation.txt",header = TRUE, sep = "", dec = ".",row.names=NULL)
full_cell_info<-data.frame(paste(mol_hs$chip_id,".",mol_hs$row.names,".",mol_hs$experiment,sep = ""),mol_hs$mol_hs)
colnames(full_cell_info)<-c("id", "mol_hs")
```


```{r,echo=FALSE}
library(dplyr)
data<-read.table("D:/Uchicago/Thesis/Real_data/scqtl-counts/scqtl-counts.txt", header = TRUE, sep = "", dec = ".")
combined_filter<-read.table("D:/Uchicago/Thesis/Real_data/combined_filter.txt", header = FALSE, sep = ",", dec = ".")
combined_filter[,1]<-combined_filter[,1]+2 #adjusting index from python to R
clean_data<-data.frame(data[,1],data[,combined_filter[,1]])
colnames(clean_data)[1]<-"gene"
gene_filter<-read.table("D:/Uchicago/Thesis/Real_data/mean/mean.txt", header = TRUE, sep = "", dec = ".",row.names=NULL)
gene_filter<-gene_filter[,c("gene","p_beta")]
clean_data<- left_join(clean_data, gene_filter, by = "gene")
clean_data <- clean_data[order(clean_data$p_beta),]
clean_data <- clean_data[,1:(ncol(clean_data)-1)]
```


```{r}
estimate_MLE_one_gene_with_si<-function(full_data,full_si){
  data_ind<-substr(colnames(full_data),0,7)
  data_ind<-data_ind[2:length(data_ind)]
  est=matrix(, nrow = (length(unique(data_ind))-1), ncol = 3)
  info=matrix(, nrow = (length(unique(data_ind))-1), ncol = 2)
  iter=0
for (i in unique(data_ind)){
  if (i=="NA18498"){
    next}
  iter=iter+1
  x=full_data[ , grepl(i, names(full_data))]
  si=data.frame(colnames(x))
  colnames(si)<-c("id")
  si<-left_join(si, full_si, by = "id")
  si=data.matrix(si[,2])
  x=t(data.matrix(x))
  est[iter,]=MaximumLikelihood_with_si(x,si)
  info[iter,1]=i
  info[iter,2]=dim(x)[1]
}
return(list(est,info))
  }
```


```{r}
options(warn=-1)
top_gene<-clean_data[1,]
results=estimate_MLE_one_gene_with_si(top_gene,full_cell_info)
est=results[[1]]
info=results[[2]]
hist(log(est[,1]),breaks=30, main="MLE estimate of k_on parameter,1000 runs, true k_on=0.1, log scale", xlab="k_on MLE estimate")
hist(log(est[,2]),breaks=30, main="MLE estimate of k_off parameter,1000 runs, true k_off=0.1, log scale", xlab="k_off MLE estimate")
hist(log(est[,3]),breaks=30, main="MLE estimate of k_r parameter,1000 runs,true k_r=100, log scale", xlab="k_r MLE estimate")
```


## Distribution for one individuals

```{r}
par(mfrow=c(1,3))
na18501=top_gene[ , grepl("NA18501", names(top_gene))]
hist(data.matrix(na18501), xlab="Number of molecules", main="Distribution of protein expression") #Real data
beta=rbeta(200,est[24,1],est[24,2])
hist(beta, main="Beta distribution")
poisson=rpois(200,beta*est[24,3])
hist(poisson, xlab="Number of molecules",main="Poisson distribution")
```

## Boxplots by genotype

As in previous vignette  we compare genotype to the estimated $k_{on}$, $k_{off}$ and $k_{r}$ parameters.

```{r,echo=FALSE}
mean_txt_file<-read.table("D:/Uchicago/Thesis/Real_data/mean/mean.txt", header = TRUE, sep = "", dec = ".",row.names=NULL)
mean_txt_file<-mean_txt_file[,c("gene","id")]
#data[1,1]
#mean_txt_file[mean_txt_file$gene == "ENSG00000113558", ]

genotypes<-read.table("D:/Uchicago/Thesis/Real_data/genotypes.txt", header = F, sep = "")
genotypes<-t(genotypes)
colnames(genotypes)<-c("ind","genotype") 

dataset<-data.frame(est,info)
colnames(dataset)<-c("k_on","k_off","k_r","ind","n_cells")
dataset<-left_join(dataset,data.frame(genotypes), by = "ind")
dataset$genotype<-as.numeric(as.character(dataset$genotype))
```


```{r, echo=FALSE}
par(mfrow=c(1,3))
boxplot(dataset[dataset$genotype=="0",1],dataset[dataset$genotype=="1",1],dataset[dataset$genotype=="2",1], main= "k_on boxplot by genotype", names=c("0", "1","2"))
boxplot(dataset[dataset$genotype=="0",2],dataset[dataset$genotype=="1",2],dataset[dataset$genotype=="2",2], main= "k_off boxplot by genotype", names=c("0", "1","2"))
boxplot(dataset[dataset$genotype=="0",3],dataset[dataset$genotype=="1",3],dataset[dataset$genotype=="2",3], main= "k_r boxplot by genotype", names=c("0", "1","2"))
```

# Regression

```{r, echo=FALSE}
par(mfrow=c(1,3))
model1<- lm(k_on~dataset$genotype, data=dataset)
model2<- lm(k_off~dataset$genotype, data=dataset)
model3<- lm(k_r~dataset$genotype, data=dataset)
boxplot(dataset[dataset$genotype=="0",1],dataset[dataset$genotype=="1",1],dataset[dataset$genotype=="2",1], main= "k_on boxplot by genotype", names=c("0", "1","2"))
abline(model1, col="red")
boxplot(dataset[dataset$genotype=="0",2],dataset[dataset$genotype=="1",2],dataset[dataset$genotype=="2",2], main= "k_off boxplot by genotype", names=c("0", "1","2"))
abline(model2, col="red")
boxplot(dataset[dataset$genotype=="0",3],dataset[dataset$genotype=="1",3],dataset[dataset$genotype=="2",3], main= "k_r boxplot by genotype", names=c("0", "1","2"))
abline(model3, col="red")
```

