---
title: "MLE_bootstrap_variability_by_sample_size"
author: "Margarita Orlova"
date: "December 19, 2019"
output:
  word_document: default
  html_document: default
---

```{r, echo=FALSE}
library(reticulate)
library(plotrix)
library(vioplot)
source_python("C:/Users/Moonkin/Documents/GitHub/Thesis_single_RNA/analysis/MLE.py")
```

This section is the continuation of bootstrap MLE estimates variability exploration. This time we would want to know the MLE bootstrap variability as function of sample size.

For each of the $k_on$, $k_off$, $k_r$ case we generate a dataset of 100,200,500 and 1000 cells and bootstrap the data 500 times to estimate parameters with MLE. Then we take a look at boxplot of MLE estimates and the standart deviation by sample size.  
All plot below are in log scale.

```{r}
bootstrap<-function(n_datasets,k_on,k_off,kr,n_cells){
  est=matrix(, nrow = n_datasets, ncol = 3)
  x=generate_data(k_on,k_off,kr,n_cells)
  for (i in 1:n_datasets){
    x=generate_data(k_on,k_off,kr,n_cells)
    est[i,]=MaximumLikelihood(x)}
  return(est)}

MLE_func<-function(n_datasets,k_on,k_off,kr,n_cells){
  MLE<-list()
  for (i in 1:length(n_cells)){
    est<-bootstrap(n_datasets,k_on,k_off,kr,as.integer(n_cells[i]))
    MLE[[i]] <- est}
  return(MLE)}
```


## I) Small $k_{on}$ and $k_{off}$ case:

In the case of small $k_{on}$ and $k_{off}$, as we saw before, parameters are identifiable so the standart deviation is quite small even for small 100 cell dataset and diminish with the increase of sample size. It is true for $k_{on}$,$k_{off}$ and $k_{r}$ as well.

```{r}
#True k_on=k_off=0.1
n_datasets=500
k_on=as.numeric(0.1)
k_off=as.numeric(0.1)
kr=as.numeric(100)
n_cells<-c(100,200,500,1000)
MLE<-MLE_func(n_datasets,k_on,k_off,kr,n_cells)
x1 <- MLE[[1]]
x2 <- MLE[[2]]
x3 <- MLE[[3]]
x4 <- MLE[[4]]
boxplot(log(x1[,1]),log(x2[,1]),log(x3[,1]),log(x4[,1]),xlab="Number of cells",names=c("100", "200","500", "1000"),main="k_on MLE bootstrap, true k_on=0.1")
abline(h = log(0.1),col="red")
paste("K_on standart deviations:","n=100:",round(sqrt(var(x1[,1],na.rm=TRUE)),3), "; n=200:",round(sqrt(var(x2[,1],na.rm=TRUE)),3), "; n=500:", round(sqrt(var(x3[,1],na.rm=TRUE)),3), "; n=1000:",round(sqrt(var(x4[,1],na.rm=TRUE)),3))

boxplot(log(x1[,2]),log(x2[,2]),log(x3[,2]),log(x4[,2]),xlab="Number of cells",names=c("100", "200","500", "1000"),main="k_off MLE bootstrap, true k_off=0.1")
abline(h = log(0.1),col="red")
paste("k_off standart deviations:", "n=100:", round(sqrt(var(x1[,2],na.rm=TRUE)),3), "; n=200:",round(sqrt(var(x2[,2],na.rm=TRUE)),3), "; n=500:", round(sqrt(var(x3[,2],na.rm=TRUE)),3), "; n=1000: ", round(sqrt(var(x4[,2],na.rm=TRUE)),3))

boxplot(log(x1[,3]),log(x2[,3]),log(x3[,3]),log(x4[,3]),xlab="Number of cells",names=c("100", "200","500", "1000"),main="k_r MLE bootstrap, true k_r=100")
abline(h = log(100),col="red")
paste("k_r standart deviations:", "n=100:", round(sqrt(var(x1[,3],na.rm=TRUE)),3), "; n=200:",round(sqrt(var(x2[,3],na.rm=TRUE)),3), "; n=500:", round(sqrt(var(x3[,3],na.rm=TRUE)),3), "n=1000:", round(sqrt(var(x4[,3],na.rm=TRUE)),3))
```


## II) Large $k_{on}$ and $k_{off}$ case: 

For large $k_{on}$ and $k_{off}$ case, we see decrease in standart deviation, but the magnitude is quite high.

```{r,echo=FALSE}
#True k_on=k_off=50
k_on=as.numeric(50)
k_off=as.numeric(50)
kr=as.numeric(100)
MLE<-MLE_func(n_datasets,k_on,k_off,kr,n_cells)
x1 <- MLE[[1]]
x2 <- MLE[[2]]
x3 <- MLE[[3]]
x4 <- MLE[[4]]
boxplot(log(x1[,1]),log(x2[,1]),log(x3[,1]),log(x4[,1]),xlab="Number of cells",names=c("100", "200","500", "1000"),main="k_on MLE bootstrap, true k_on=50")
abline(h = log(50),col="red")
paste("K_on standart deviations:","n=100:",round(sqrt(var(x1[,1],na.rm=TRUE)),3), "; n=200:",round(sqrt(var(x2[,1],na.rm=TRUE)),3), "; n=500:", round(sqrt(var(x3[,1],na.rm=TRUE)),3), "; n=1000:",round(sqrt(var(x4[,1],na.rm=TRUE)),3))

boxplot(log(x1[,2]),log(x2[,2]),log(x3[,2]),log(x4[,2]),xlab="Number of cells",names=c("100", "200","500", "1000"),main="k_off MLE bootstrap, true k_off=50")
abline(h = log(50),col="red")
paste("k_off standart deviations:", "n=100:", round(sqrt(var(x1[,2],na.rm=TRUE)),3), "; n=200:",round(sqrt(var(x2[,2],na.rm=TRUE)),3), "; n=500:", round(sqrt(var(x3[,2],na.rm=TRUE)),3), "; n=1000: ", round(sqrt(var(x4[,2],na.rm=TRUE)),3))

boxplot(log(x1[,3]),log(x2[,3]),log(x3[,3]),log(x4[,3]),xlab="Number of cells",names=c("100", "200","500", "1000"),main="k_r MLE bootstrap, true k_r=100")
abline(h = log(100),col="red")
paste("k_r standart deviations:", "n=100:", round(sqrt(var(x1[,3],na.rm=TRUE)),3), "; n=200:",round(sqrt(var(x2[,3],na.rm=TRUE)),3), "; n=500:", round(sqrt(var(x3[,3],na.rm=TRUE)),3), "n=1000:", round(sqrt(var(x4[,3],na.rm=TRUE)),3))
```


## III) Large value of $k_{off}$ and small $k_{on}$ case:

For small $k_{on}$ large $k_{off}$ we observe that as $k_{on}$ has small deviation and as we might remember from previous vignette, this is the only identifiable parameter in this scenario. Both  $k_{off}$ and $k_{r}$ were unidentifiable, so we see high standart deviation even with big data size. We also experience some instability in standart deviation - sample size dependance as we sometimes see increase in standart deviation despite the sample size growth.

```{r, echo=FALSE}
#True k_on=1, k_off=10
k_on=as.numeric(1)
k_off=as.numeric(10)
kr=as.numeric(100)
MLE<-MLE_func(n_datasets,k_on,k_off,kr,n_cells)
x1 <- MLE[[1]]
x2 <- MLE[[2]]
x3 <- MLE[[3]]
x4 <- MLE[[4]]
boxplot(log(x1[,1]),log(x2[,1]),log(x3[,1]),log(x4[,1]),xlab="Number of cells",names=c("100", "200","500", "1000"),main="k_on MLE bootstrap, true k_on=1")
abline(h = log(1),col="red")
paste("K_on standart deviations:","n=100:",round(sqrt(var(x1[,1],na.rm=TRUE)),3), "; n=200:",round(sqrt(var(x2[,1],na.rm=TRUE)),3), "; n=500:", round(sqrt(var(x3[,1],na.rm=TRUE)),3), "; n=1000:",round(sqrt(var(x4[,1],na.rm=TRUE)),3))

boxplot(log(x1[,2]),log(x2[,2]),log(x3[,2]),log(x4[,2]),xlab="Number of cells",names=c("100", "200","500", "1000"),main="k_off MLE bootstrap, true k_off=10")
abline(h = log(10),col="red")
paste("k_off standart deviations:", "n=100:", round(sqrt(var(x1[,2],na.rm=TRUE)),3), "; n=200:",round(sqrt(var(x2[,2],na.rm=TRUE)),3), "; n=500:", round(sqrt(var(x3[,2],na.rm=TRUE)),3), "; n=1000: ", round(sqrt(var(x4[,2],na.rm=TRUE)),3))

boxplot(log(x1[,3]),log(x2[,3]),log(x3[,3]),log(x4[,3]),xlab="Number of cells",names=c("100", "200","500", "1000"),main="k_r MLE bootstrap, true k_r=100")
abline(h = log(100),col="red")
paste("k_r standart deviations:", "n=100:", round(sqrt(var(x1[,3],na.rm=TRUE)),3), "; n=200:",round(sqrt(var(x2[,3],na.rm=TRUE)),3), "; n=500:", round(sqrt(var(x3[,3],na.rm=TRUE)),3), "n=1000:", round(sqrt(var(x4[,3],na.rm=TRUE)),3))
```


## IV) Small value of $k_{off}$ and big $k_{on}$

Small value if $k_{off}$ and big $k_{on}$ scenario has identifiable parameters, so alike with small value $k_{on}$, $k_{on}$ case we observe reasonable standart deviation values, especially for bigger dataset as well as decrease in standart deviation size. 

```{r, echo=FALSE}
#True k_on=10, k_off=1
k_on=as.numeric(10)
k_off=as.numeric(1)
kr=as.numeric(100)
MLE<-MLE_func(n_datasets,k_on,k_off,kr,n_cells)
x1 <- MLE[[1]]
x2 <- MLE[[2]]
x3 <- MLE[[3]]
x4 <- MLE[[4]]
boxplot(log(x1[,1]),log(x2[,1]),log(x3[,1]),log(x4[,1]),xlab="Number of cells",names=c("100", "200","500", "1000"),main="k_on MLE bootstrap, true k_on=10")
abline(h = log(10),col="red")
paste("K_on standart deviations:","n=100:",round(sqrt(var(x1[,1],na.rm=TRUE)),3), "; n=200:",round(sqrt(var(x2[,1],na.rm=TRUE)),3), "; n=500:", round(sqrt(var(x3[,1],na.rm=TRUE)),3), "; n=1000:",round(sqrt(var(x4[,1],na.rm=TRUE)),3))

boxplot(log(x1[,2]),log(x2[,2]),log(x3[,2]),log(x4[,2]),xlab="Number of cells",names=c("100", "200","500", "1000"),main="k_off MLE bootstrap, true k_off=1")
abline(h = log(1),col="red")
paste("k_off standart deviations:", "n=100:", round(sqrt(var(x1[,2],na.rm=TRUE)),3), "; n=200:",round(sqrt(var(x2[,2],na.rm=TRUE)),3), "; n=500:", round(sqrt(var(x3[,2],na.rm=TRUE)),3), "; n=1000: ", round(sqrt(var(x4[,2],na.rm=TRUE)),3))

boxplot(log(x1[,3]),log(x2[,3]),log(x3[,3]),log(x4[,3]),xlab="Number of cells",names=c("100", "200","500", "1000"),main="k_r MLE bootstrap, true k_r=100")
abline(h = log(100),col="red")
paste("k_r standart deviations:", "n=100:", round(sqrt(var(x1[,3],na.rm=TRUE)),3), "; n=200:",round(sqrt(var(x2[,3],na.rm=TRUE)),3), "; n=500:", round(sqrt(var(x3[,3],na.rm=TRUE)),3), "n=1000:", round(sqrt(var(x4[,3],na.rm=TRUE)),3))

```




