---
title: "Real_data_filter"
author: "Margarita Orlova"
date: "January 28, 2020"
output: html_document
---

In this vignette we will load the data, pre-process it and analyse top  gene by p_beta parameter.

## Filtering

```{r}
library(dplyr)
data<-read.table("D:/Uchicago/Thesis/Real_data/scqtl-counts/scqtl-counts.txt", header = TRUE, sep = "", dec = ".")
```

```{r}
combined_filter<-read.table("D:/Uchicago/Thesis/Real_data/combined_filter.txt", header = FALSE, sep = ",", dec = ".")
combined_filter[,1]<-combined_filter[,1]+2 #adjusting index from python to R
clean_data<-data.frame(data[,1],data[,combined_filter[,1]])
colnames(clean_data)[1]<-"gene"
gene_filter<-read.table("D:/Uchicago/Thesis/Real_data/mean/mean.txt", header = TRUE, sep = "", dec = ".",row.names=NULL)
gene_filter<-gene_filter[,c("gene","p_beta")]
clean_data<- left_join(clean_data, gene_filter, by = "gene")
clean_data <- clean_data[order(clean_data$p_beta),]
clean_data <- clean_data[,1:(ncol(clean_data)-1)]
```


## Maximum likelihood estimation for one gene

```{r}
library(reticulate)
library(robustbase)
source_python("C:/Users/Moonkin/Documents/GitHub/Thesis_single_RNA/analysis/MLE.py")
```


```{r}
estimate_MLE_one_gene<-function(full_data){
  data_ind<-substr(colnames(full_data),0,7)
  data_ind<-data_ind[2:length(data_ind)]
  est=matrix(, nrow = (length(unique(data_ind))-1), ncol = 3)
  info=matrix(, nrow = (length(unique(data_ind))-1), ncol = 2)
  iter=0
for (i in unique(data_ind)){
  if (i=="NA18498"){
    next}
  iter=iter+1
  x=full_data[ , grepl(i, names(full_data))]
  x=t(data.matrix(x))
  est[iter,]=MaximumLikelihood(x)
  info[iter,1]=i
  info[iter,2]=dim(x)[1]}
return(list(est,info))}
```


```{r}
top_gene<-clean_data[1,]
x=estimate_MLE_one_gene(top_gene)
```


```{r}
estimates=x[[1]]
par(mfrow=c(1,3))
hist(log(estimates[,1]), main="Distribution of k_on estimates", xlab="")
hist(log(estimates[,2]), main="Distribution of k_off estimates", xlab="")
hist(log(estimates[,3]), main="Distribution of k_r estimates", xlab="")
```


## Distribution for one individuals

Then we take the individual with most cells (NA18501) and take a look at the estimated parameters:

The first plot is the real protein level distribution for NA18501 individual.
Second and third plots are beta and poisson distributions with the estimated $k_{on}$,$k_{off}$ and $k_r$ parameters.


```{r}
par(mfrow=c(1,3))
na18501=top_gene[ , grepl("NA18501", names(top_gene))]
hist(data.matrix(na18501), xlab="Number of molecules", main="Distribution of protein expression") #Real data
beta=rbeta(1000,estimates[24,1],estimates[24,2])
hist(beta)
poisson=rpois(1000,beta*estimates[24,3])
hist(poisson, xlab="Number of molecules")
```









